A Domain Specific Language for SpectreGuard

### 2020-12-11 paper/main.tex
Intro; upcoming pattern matching paper citation

Haskell is a purely functional programming language with strong static type system
based on System F and algebraic datatypes (ADTs).~\cite{HistoryOfHaskell} These features,
particularly the type system and the algebraic datatypes, make it a powerful tool
for the creation of EDSLs. Expressions, commands and programs (compound commands)
can be straightforwardly expressed using ADTs using a \textit{deep embedding}.~\cite{DeepAndShallow}

@@

This also outlines procedure that the C code generator uses. When code written
in the EDSL is transformed into C code, a name generator first annotates the
abstract syntax tree (AST) with globally unique names (enforcing the Barendregt
variable convention; see Chapter 2 in~\cite{BarendregtConversion}). This annotated AST can be used by both the code generator and the
information flow leak analysis tool (the latter is described in Section
\ref{sec:Analysis}).

The one major aspect of the C code generator that is not described in this table
is how the loop variable in \ttt{for} loops can be tagged as secret (if appropriate).
This is accomplished by recovering the type-level sensitivity type information at
runtime. Such a recovery is made possible by the use of singletons.~\cite{Singletons}

@@

Singletons~\cite{SingletonsPaper} are used to preserve this type level at runtime, so that the
code generator can generate SpectreGuard annotations where appropriate. These annotations can
then be used by the SpectreGuard mitigation system.~\cite{SpectreGuard}

@@

These operations are used by the tool in its analysis of an EDSL
program as follows:

@@

\section{Future Work}
\label{sec:FutureWork}
A further investigation into the tradeoffs and possible additional features of
the information flow analysis is likely to be useful. This could provide a more
accurate and succinct picture to programmers of where leaks are occurring, or
potentially occurring. It is possible that some of these other leak conditions
could be incorporated into the type system, by exploiting the dependent type
features enabled by singletons~\cite{SingletonsPaper} and, in the future, by
Dependent Haskell.~\cite{DepHaskSpec} Dependent types provide a powerful
technique for writing programs at a type-level, which will be executed at
compile-time by the type checker.~\cite{CertProg, DepHaskSpec} A more
lightweight technique, which might be able to be applied to allow for more
expressive policies to control of information flow within the type system, would
be the ``ghosts of departed proofs'' design technique.~\cite{DepartedProofs}

@@

Additionally, features to safely convert between sensitivity levels could prove
useful. A conversion from public to secret would be straightforward. A situational conversion
going the other direction could be provided by an instruction in the EDSL
which might perform a Spectre mitigation measure, such as a cache flush. \cite{PLtea-james}

More primitive operations and support for a foreign function interface (FFI)
with C would make the EDSL more practical. This would be straightforward to add,
as the system already closely resembles C code and generates C code.

@@

SpectreGuard is a powerful tool for protecting code against Spectre attacks and,
in the present work, we have given tools which provide additional guarantees
that this tool will function as intended.

>>>>

Haskell is a purely functional programming language with strong static type
system based on System F and algebraic datatypes (ADTs).~\cite{HistoryOfHaskell}
These features, particularly the type system, the algebraic datatypes and its
support for higher-order functions, make it a powerful tool for the creation of
EDSLs. Expressions, commands and programs (compound commands) can be
straightforwardly expressed using ADTs using a \textit{deep
embedding}.~\cite{DeepAndShallow}

@@

This also outlines procedure that the C code generator uses. When code written
in the EDSL is transformed into C code, a name generator first annotates the
abstract syntax tree (AST) with globally unique names (enforcing the Barendregt
variable convention; see Chapter 2 in~\cite{BarendregtConversion}). This
annotated AST can be used by both the code generator and the information flow
leak analysis tool (the latter is described in Section \ref{sec:Analysis}). It
is worth noting that, as a result of this technique, the names in the generated
C code will not generally be the same as the names in the EDSL code.

The one major aspect of the C code generator that is not described in this table
is how the loop variable in \ttt{for} loops can be tagged as secret (if appropriate).
This is accomplished by recovering the type-level sensitivity type information at
runtime. Such a recovery is made possible by the use of singletons.~\cite{Singletons}
\ttt{for} loops provide a particular example of where Haskell's support for higher-order function
can be used inside an EDSL to implement commands that involve the creation of
new variables.

@@

Singletons~\cite{Singletons} are used to preserve this type level at runtime, so that the
code generator can generate SpectreGuard annotations where appropriate. These annotations can
then be used by the SpectreGuard mitigation system.~\cite{SpectreGuard}

@@

\noindent These operations are used by the tool in its analysis of an EDSL
program as follows:

@@

\section{Future Work}
\label{sec:FutureWork}
A further investigation into the tradeoffs and possible additional features of
the information flow analysis is likely to be useful. This could provide a more
accurate and succinct picture to programmers of where leaks are occurring, or
potentially occurring. It is possible that some of these other leak conditions
could be incorporated into the type system, by exploiting the dependent type
features enabled by singletons~\cite{Singletons} and, in the future, by
Dependent Haskell.~\cite{DepHaskSpec} Dependent types provide a powerful
technique for writing programs at a type-level, which will be executed at
compile-time by the type checker.~\cite{CertProg, DepHaskSpec} A more
lightweight technique, which might be able to be applied to allow for more
expressive policies to control of information flow within the type system, would
be the ``ghosts of departed proofs'' design technique.~\cite{DepartedProofs}

@@

Additionally, features to safely convert between sensitivity levels could prove
useful. A conversion from public to secret would be straightforward. A situational conversion
going the other direction could be provided by an instruction in the EDSL
which might perform a Spectre mitigation measure, such as a cache flush.~\cite{PLtea-james}

More primitive operations and support for a foreign function interface (FFI)
with C would make the EDSL more practical. This would be straightforward to add,
as the system already closely resembles C code and generates C code. This could allow
for better integration between code written in the EDSL and regular C code. One potential
practical application of this is that the security-critical portion of a program
can be written in this EDSL and the rest of the application can be written as a regular
C program which interacts with the security-critical portion via an FFI going from C to
the EDSL and an FFI going the other direction.

Additional primitive operations added to the EDSL could include higher-level language features, such as higher-order
functions, algebraic data types, pattern matching and tail recursion. An implementation path for
these features is described in the upcoming paper \textit{On Adding Pattern Matching to Haskell-based Deeply Embedded Domain Specific Languages}, by
David Young, Mark Grebe and Andy Gill.~\cite{PatMatchingEDSL}

@@

SpectreGuard is a powerful mechanism for protecting code against Spectre attacks and,
in the present work, we have given tools which provide additional guarantees
that this mechanism will function as intended.
